Periodicity of Geodesic Flows on the Necker Cube Surface

### 2017-11-13 newpaper.tex
110617
111217

\documentclass[]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{svg}
\usepackage{enumitem}
\usepackage{color}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}

@@

\def\bU{\mathbf{U}} \def\btU{\tilde{\bU}} \def\bUs{\bU^\circ}
\def\btUos{\btU_0^\circ}
\def\bG{\mathbf{G}} \def\bGs{\mathbf{G}^\circ}
\def\hol{\mathbf{hol}}
\def\bM{\mathbf{M}}
\def\bMs{\mathbf{M}^\circ}

@@

\section{Periodic Tiling of Necker Cubes}
This section will detail how the Necker Cube surface is obtained as an infinite regular cover of a branched torus. Consider the Necker cube with the following identifications:
\begin{figure}[H]
\centering
\includesvg[width=2in.]{neckercube}
\end{figure}

We denote this surface $\bG$. $\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a unit cube. $\bG$ is a genus one surface with three vertices homeomorphic to $\TT^2$. Check that every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$. These are the conical singularities of the surface, $\Sigma$. We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$. Thus, $\pi_1(\bGs)\cong\pi_1((thrice\text{-}branched~torus))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators.
\begin{Def}
The \textbf{Necker cube surface} $\mathbf{S}$ is obtained as the regular cover of $\bG$ by periodically tiling along the edges with solid triangles.\\
The \textbf{branched Necker cube surface} $\mathbf{S}^\circ$ is obtained similarly by tiling $\bGs$.
\end{Def}

\noindent Consider the following families of unit squares in $\RR^3$:\\
\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}

With the proper identifications on  $\mathbf{A}_{0,0,0}\cup\mathbf{B}_{0,0,0}\cup\mathbf{C}_{0,0,0}$, we have $\bG$'s embedded form in $\RR^3$ and a map $s:\bG\hookrightarrow\RR^3$. Similarly, removing the integer points from this set is an embedding of $\bGs$. Check that tiling along the edges as previously described determines a set of points in $\RR^3$ that coincides with points on $\mathbf{S}$ with $s^*$ as an extension of $s$ to the cover. That is, $s(\mathbf{S})=\bigcup\left\{\mathbf{A}_{m,n,p}\cup\mathbf{B}_{m,n,p}\cup\mathbf{C}_{m,n,p}~:~m+n+p=0\right\}$. Thus a point $x_0\in\mathbf{S}$ is in one-to-one correspondence with this set. To avoid ambiguity we say $U$ is a \emph{neighborhood} of $s(x_0)$ if it is contained in exactly one of $\mathbf{A}_{m,n,p}, \mathbf{B}_{m,n,p}, \mathbf{C}_{m,n,p}$. As $U$ must contain an open ball contained on a single square there cannot be a neighborhood where $s(x_0)\in\ZZ^3$ and $x_0$ is a conical point. It follows then that a neighborhood in $\mathbf{S}$ is a neighborhood in $\mathbf{S}^\circ$ as well.

\subsection{Translation Surface Cover of $\bGs$}
Reconstruct $\bG$ and $\bGs$ to resemble a quotient of the flattened Necker cube surface by a series of cutting and gluing operations: \\
\begin{figure}[H]
\centering
\includesvg[width=3in]{LTorus}
\end{figure}
\noindent This L-shaped torus is then cut along the dotted lines and pieced together to look like a $2\times2$ torus with a unit square removed from the center:\\
\begin{figure}[H]
\centering
\includesvg[width=2in]{BranchTorus}
\caption{$\bGs$ with 4 independent, closed paths labeled identified}
\end{figure}
The paths labeled $a,b,c,d$ are independent. Let $\pi_1(\bGs)=\<a,b,c,d\>$. $\bGs$ has a $\ZZ^2$ tiling by traversing along paths $a,b$ and attaching infinitely many copies of the surface. It's apparent that this $\ZZ^2$ tiling is the flattened Necker cube surface. We use group homomorphisms on this free group to construct translation surfaces and $\ZZ^2$-covers to better understand geodesic behavior on the Necker cube surface.

\begin{Def}
Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$
\end{Def}

@@

We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy. Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$. Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ to be in $\RR^2$. Let $x_0\in\bGs$. The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$. Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$. We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$. An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$. Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$. Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ the monodromy group is discrete and well-defined on $\bGs$.

@@

\begin{Def}
The surface $\bUs$ is the regular cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi_1$. 
\end{Def}

The deck group of this cover is isomorphic to $\ZZ^2$. By reversing the cutting and gluing done previously to $\bGs$ and $\bG$, one can easily see that a cover defined as the kernel of this homomorphism is identical to the tiling of Necker cubes along two of its edges. For this reason, $\bU\cong\mathbf{S}$.

@@

\begin{figure}[H]
\centering
\includegraphics[width=4in.]{closed2.png}
\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
\label{fig:complicated}
\end{figure}

$\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
\subsection{Veech Staircase Model}
Let $f:\btU\rightarrow\mathbf{M}$ be a projection onto the quotient surface that takes every point to its modular equivalent by its $\ZZ^2$ symmetries, and what we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$

\begin{figure}[H]
\centering
\includesvg[width=2.6in]{mtilda}
\caption{Compact translation surface, $\mathbf{M}$, covered by the infinite surface with edges and cone singularities (1,2,3,4) identified. The Roman numerals are meant to identify each quotient with a plane in the cover, $\mathbf{\tilde{\mathbf{U}}}$.}
\label{fig:mtilda}
\end{figure}

@@

>>>>

\documentclass[]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{svg}
\usepackage{enumitem}
\usepackage{color}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage[arrow]{xy}
\usepackage{tikz-cd}

@@

\def\bU{\mathbf{U}} \def\btU{\tilde{\bU}} \def\bUs{\bU^\circ}
\def\btUos{\btU_0^\circ}
\def\bG{\mathbf{G}} \def\bGs{\mathbf{G}^\circ}
\def\hol{\mathbf{hol}}
\def\bM{\mathbf{M}}
\def\bMs{\mathbf{M}^\circ}
\def\btUs{\btU^\circ} 
\def\xtild{\tilde{x}_0}
\def\utild{\tilde{u}_0}
\def\gamtild{\tilde{\gamma}}

@@

\section{Periodic Tiling of Necker Cubes}
This section will detail how the Necker Cube surface is obtained as an infinite regular cover of a branched torus. Consider the Necker cube with the following identifications:
\begin{figure}[H]
\centering
\includesvg[width=1.6in.]{neckercube}\includesvg[width=2in.]{neckercubepath}
\end{figure}

We denote this surface $\bG$. $\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a unit cube. $\bG$ is a genus one surface with three vertices homeomorphic to $\TT^2$. Check that every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$. These are the conical singularities of the surface, $\Sigma$. We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$. Thus, $\pi_1(\bGs)\cong\pi_1((\TT^2 \text{ with 3 punctures}))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators. We may flatten $\bG$ and $\bGs$ into an L-shaped torus: \\
\begin{figure}[H]
\centering
\includesvg[width=3in]{LTorus}
\end{figure}
\noindent This L-shaped torus is then cut along the dotted lines and pieced together to resemble a $2\times2$ torus with a unit square removed from the center:\\
\begin{figure}[H]
\centering
\includesvg[width=2in]{BranchTorus}
\caption{$\bGs$ with 4 independent, closed paths labeled identified}
\end{figure}
Denote its fundamental group by $\pi_1(\bGs)=\<a,b,c,d\>$, where $a,b,c,d$ are the labeled independent paths on the surface. Consider the following map:

\begin{Def}
Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$
\end{Def}

\begin{Def}
The punctured \textbf{Necker Cube Surface} is a regular cover of $\bGs$, and is denoted $\bUs$. It's fundamental group presentation is the subgroup of $\pi_1(\bGs)$, $\pi_1(\bUs)=\ker\varphi_1$.
\end{Def}

Thus $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong \varphi_1(\pi_1(\bGs))=\ZZ^2$ is the associated \emph{deck group} of the cover. $\bUs$ is the primary object of study.

\subsection{Monodromy Group Representation}
In this section we describe a translation cover of $\bGs$ which and $\bUs$ which is factored through as the kernel of a homomorphism from $\pi_1(\bGs)\rightarrow \SO(2,\ZZ)$. This homomorphism represents the monodromy group as the fundamental group acting on the unit tangent bundles $T^1\bGs$ and $T^1\bUs$.

@@

We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy. Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$. Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ as vectors in $\RR^2$. Let $x_0\in\bGs$. The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$. Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$. We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$. An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$. Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$. Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ monodromy group is discrete and well-defined on $\bGs$.

@@

\begin{Def}
Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
\end{Def}

\compav{Prove that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$}

$\btUs$ is also a cover of $\bMs$ with fundamental group $\ker\varphi_1\big\vert_{\pi_1(\bMs)}$. In the former case the deck group of the cover is $\SO(2,\ZZ)$, while in the latter case it is $\ZZ^2$. The deck group of the cover $\btUs\rightarrow\bGs$ is $\ZZ^2\rtimes\SO(2,\ZZ)$. \compav{worth talking about?}. $\btUs$ is a four-fold cover of $\bUs$ with trivial holonomy visualized as four copies of $\bUs$, which can be thought of as countably many copies of . We can identify a closed path on $\bGs$ as an element of the deck group $\ZZ^2\times \SO(2,\ZZ)$ or, equivalently, $ 2\ZZ+2\ZZ i\times \ZZ/4\ZZ$ if we want to consider a metric on the ambient space $\CC$ that is consistent with the tiling of $\bGs$. Further, we can rotate everything in $\CC\times \ZZ/4\ZZ$ along with $\btUs$ according to the element of $\SO(2,\ZZ)$ that acts non-trivially on the holonomy of $\bUs$. For example, let $P\in2\ZZ+2\ZZ i$ and consider the following curve on $\bUs$:

\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group isomorphic to $2\ZZ+2\ZZ i \times \ZZ/4\ZZ$. Observe that it will only close if the curve acts trivially on the holonomy of $\bUs$:

@@

\subsection{Geodesics on $\bUs$}
The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ are identical in the sense that their images are in a one-to-one correspondence. Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface $\gamma:[0,\infty)\rightarrow S$ such that $\gamma'(0)=v_0\in\RR^2$ and $|| v_0||=1$. A geodesic is \emph{periodic} if for some $T\in\RR^+$ $\gamma(t)=\gamma(t+T)$ for all $t$. A geodesic is \emph{drift-periodic} if for some $T\in\RR^+$, there is some non-trivial translation of the surface that takes $\gamma(t)\mapsto \gamma(t+T)$ for all $t$. In either case we say that $T$ is the \emph{period} or \emph{length} of a segment; in the case where a geodesic does have a period, we restrict $\gamma$ to the closed interval $[0,T]$ and call this a \emph{geodesic segment}. In this paper we do not consider cases where a geodesic hits a cone singularity. For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:

\begin{figure}[H]
\centering
\includegraphics[width=4in.]{closed2.png}
\caption{Periodic geodesic path on $\bUs$ modeled with sage-flatsurf.}
\label{fig:complicated}
\end{figure}

We refer later to the following commuting diagram of branched covering maps and inclusion maps. It is meant to illustrate the relationship between all of these spaces.

\begin{equation}
\begin{tikzcd}
\arrow[two heads, bend left]{rrr}{f^\circ}[swap]{} \arrow[hook]{d}{pr_b} \btUs \arrow[two heads]{r}{r^\circ} & \bUs \arrow[hook]{d}{pr_c} \arrow[two heads]{r}{q^\circ} & \bGs \arrow[hook]{d}{pr_d} &\arrow[two heads]{l}{w^\circ} \bMs \arrow[hook]{d}{pr_a}  \\
\arrow[two heads, bend right]{rrr}{}[swap]{f} \btU \arrow[two heads]{r}{r} & \bU \arrow[two heads]{r}{q} & \bG & \arrow[two heads]{l}{w}\bM  \\
\end{tikzcd}
\label{eq:commute}
\end{equation}

Let $x_0\in G^\circ$ and fix $u_0\in q^{\circ-1}(x_0)$ Let $\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$, and $\alpha'(0)=v_0$. Denote $\alpha$'s unique lift to $\bUs$ under $q^\circ$ as $\gamma:[0,b]$, where $\gamma(0)=u_0$, $\gamma'(0)=\gamma'(b)=v_0$, and $\gamma(b)\in q^{\circ-1}(x_0)$. Fix an arbitrary point $\utild\in r^{\circ -1}(u_0)$. We define the unique lift of $\gamma$ under $r$ as the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$ and $\tilde{\gamma}'(0)=\tilde{\gamma}'(b)=v_0$. Denote their homotopy class representatives as $[\alpha],[\gamma],$ and $[\tilde{\gamma}]$.

A geodesic must have trivial holonomy upon returning to its initial point on $\bGs$. If a geodesic segment does not lift to a closed curve on $\bUs$, then $\gamma(b)$ would be a translation of $u_0$ by some element of the deck group of the cover, i.e. $\varphi_1([\alpha])\in\ZZ^2$ is non-trivial, but $\varphi_2( [ \alpha] ) =I_2$ as $\varphi_2$ is a monodromy map whose image acts on $Hol_p (\triangledown)$ of $\bUs$. Therefore $\gamma'(b)=v_0$ as well. We can naturally extend these segments to the domain $[0,\infty)$ by continuously concatenating the segments.

\begin{Def}
Let $m\in\NN$. The concatenation of a geodesic segment lifted from $\alpha$ onto $\bUs$ is done in the following manner:\\
\begin{equation}
\gamma^m=
\begin{cases}
\gamma(t)  & ~~~~~~~~~~0\leq  t\leq b\\
\varphi_1([\alpha])\cdot \gamma(t)  & ~~~~~~~~~~b\leq  t\leq 2b\\
\vdots\\
\varphi_1([\alpha]^{m-1})\cdot \gamma(t)  & (m-1)b\leq t\leq mb\\
\end{cases}
\label{eq:concat}
\end{equation}
 \\
The new segment is a map $\gamma^m:[0,mb]\rightarrow\bUs$ that is well-defined on the interval. Let $\Phi=\lim\limits_{m\to\infty} \gamma^m$. We call $\Phi_t:[0,\infty)\rightarrow\bUs$ a \textbf{geodesic} on the Necker cube surface.
\end{Def}

\begin{Def}
The geodesic $\tilde{\Phi}_t:[0,\infty)\rightarrow\btUs$ is a geodesic obtained in a similar manner to $\eqref{eq:concat}$ by lifting $\gamma$ under $r^\circ$ and continuously concatenating $\tilde{\gamma}$ with the exception that the element of the deck group acting after every iteration is $(\varphi_2([\alpha]),\varphi_1([\alpha]))\in\SO(2,\ZZ)\times \ZZ^2$. 
\end{Def}

The geodesic $\Phi$ obtained in this manner has a period that depends entirely on the length of the closed segment on $\bGs$. In addition, $\Phi(t+mb)=m\varphi_1([\alpha])\cdot\Phi(t)$ and $tilde{\Phi}(t+mb)=(m\varphi_1([\alpha]),I_2)\cdot\tilde{\Phi}(t)$  for all $t,m\geq 0$. The second statement holds since $\alpha$ and $\gamma$ must have trivial holonomy at their terminal points. We make the following observations about these lifts:

\begin{thm}
$\Phi$ is periodic if and only if $\tilde{\Phi}$ is periodic. Further, both geodesics have period $||\alpha||$.
\begin{proof}
Suppose $\Phi$ is periodic. Then $\gamma(0)=\gamma(b)=u_0$. Since $\gamma$ has trivial holonomy, $\varphi_2\big\vert_{\ker\varphi_1}([\gamma])=I_2$, and so $[\tilde{\gamma}]\in\pi_1(\btUs)$. Hence $\gamtild(0)=\gamtild(b)=\utild$. Since $\tilde{\Phi}$ is obtained by concatenating $\gamtild$ under the action of the trivial deck group element of $[\alpha]$, $\gamtild^m=\gamtild(t)$ for all $t\in[0,mb]$. It follows that $\tilde{\Phi}$ is periodic as well. Conversely, the proof is trivial since $\btUs$ is a finite degree cover of $\bUs.$ Since $||\alpha||=b$, there can be no smaller period for its lifted paths as that would imply either $\alpha$ has non-trivial holonomy or does not lift to a closed segment on $\bUs$.
\end{proof}
\end{thm}

\begin{thm}
$\Phi$ is drift-periodic if and only if $\tilde{\Phi}$ is drift-periodic. Both geodesics have period $||\alpha||$.
\begin{proof}
Suppose $\Phi$ is drift-periodic. Then $\varphi_1([\alpha])=(x,y)$ is non-trivial, but $\varphi_2([\alpha])=I_2$ is. The element of the deck group acting on $\gamtild^m(t)$ for $m>1$ is $(I_2,n(x,y))$ for $1<n<m$. As $t\to mb$ and $m\to\infty$, $d(\tilde\Phi(t),\utild)=d((I_2,m(x,y))\cdot \utild, \utild)\to\infty$, and $\tilde{\Phi}(t+b)=(I_2,(x,y))\cdot \tilde{\Phi}(t)$ is periodically translated by some element in $\ZZ^2$. Conversely, if $\tilde{\Phi}$ is drift-periodic there are three cases:\\
Case one: $\varphi_2([\alpha])$ is trivial, but $\varphi_1([\alpha])$ is not. Then $\Phi$ must be drift-periodic as well since $\gamma$ would terminate on $\varphi_1([\alpha])\cdot u_0\neq u_0$.\\
Case two: $\varphi_1([\alpha])$ is trivial, but $\varphi_2([\alpha])$ is not. This is a contradiction since that would imply $\alpha$ has non-trivial holonomy.
Case three:$\varphi_1([\alpha]),\varphi([\alpha])$ are both non-trivial. Again this is not possible since $\alpha,\gamma$ must both have non-trivial holonomy.\\
Now both geodesics must have the same period since they are obtained by concatenating unique lifts of $\alpha$.
\end{proof}
\end{thm}

These theorems essentially tell us that these kinds of geodesics behave identically on the Necker cube surface and its cover. Now we show that this is true for $\btUs$ and $\btU$. For these proofs we will refer to the induced homomorphisms from ($\ref{eq:commute}$):

\begin{equation}
\begin{tikzcd}
\arrow[hookrightarrow, bend left]{rrr}{f_*^\circ}[swap]{} \arrow[two heads]{d}{PR_b} \pi_1(\btUs) \arrow[hook]{r}{r_*^\circ} & \pi_1(\bUs) \arrow[two heads]{d}{PR_c} \arrow[hook]{r}{q_*^\circ} & \pi_1(\bGs) \arrow[two heads]{d}{PR_d} & \ar[hookrightarrow]{l}{}[swap]{w_*^\circ} \pi_1(\bMs) \arrow[two heads]{d}{PR_a} \\
\arrow[hookrightarrow, bend right]{rrr}{}[swap]{f_*}  \pi_1(\btU) \arrow[hook]{r}{r_*} & \pi_1(\bU) \arrow[hook]{r}{q_*} & \pi_1(\bG) & \ar[hookrightarrow]{l}{}[swap]{w_*} \pi_1(\bM)\\
\end{tikzcd}
\label{eq:homomorphism}
\end{equation}

\begin{Def}
Let $\tilde{\phi}:[0,\infty)\rightarrow\btUs$ be a geodesic with initial direction $v_\in$ 
\end{Def}

\begin{thm}
content...
\end{thm}

\subsection{Translation Surface}
Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries. What we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$

\begin{figure}[H]
\centering
\includesvg[width=2.6in]{mtilda}
\caption{Compact translation surface, $\mathbf{M}$ with edges and cone singularities (1,2,3,4) identified. The Roman numerals are meant to identify every ``copy" of $\bG$ with a direction differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$. The individual copies are then rotated so we have a canonical notion of direction }
\label{fig:mtilda}
\end{figure}

@@

$S=\cO\cup \cE$.

\begin{lem}
Let $\XX'$ act on $\cS$ by left matrix multiplication. Then $\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
\begin{proof}
Let $v=kv_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cS$ with $||v_0||=1$ and $k\in\RR$, and let $A=\begin{pmatrix}a & b \\ c & d
\end{pmatrix}\in\XX'$. The kernel of the homomorphism $SL(2, \ZZ)\rightarrow \SL(2,\ZZ/2\ZZ)$ that maps a matrix to its $(mod~2)$ equivalent is precisely $\XX'$ since it maps the generators trivially. [quote this paper for the proof of this maybe? http://www.fysik.su.se/~ingemar/SL2.pdf ]\\
$x,y$ are relatively prime, so Bezout's identity states there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$. It's obvious that this action is faithful on the sets since the stabilizer of any vector in $\RR^2$ is trivial. Suppose $v\in\cO$. Consider the orbit of $\begin{pmatrix}1 \\1 \end{pmatrix}$. Then there is some matrix $A\in\XX'$ such that $A\begin{pmatrix}1\\1\end{pmatrix}=v$. Since $A\in\SL(2,\ZZ)$, $\det A = ad-bc=1$. Obtain from $1=(x-b)(y-c)-bc=xy-xc-by$ that $xy-1=xc+by$. From $(*)$ we see that $xy-1=x(u(xy-1))+y(v(xy-1))$. Let $c=u(xy-1)$ and $b=v(xy-1)$. Since $xy-1$ is even, $c,b$ are even. It follows then that $a=x-b$ and $d=y-c$ are odd. Therefore $A\in \XX'$ and the action is transitive on $\cO$. If $v\in\cE$, then either $x$ or $y$ is even. In either case this would mean that either $a$ or $d$ is even, which would mean $A\notin\XX'$, a contradiction.\\
Now consider the orbit of $\begin{pmatrix}1\\0\end{pmatrix}$. Left multiplying by $A$ would mean $a=x$ and $c=y$. Split $\cE$ into sets $\cE_1=\{(x,y)\in\cE :x\in2\ZZ,~y\in(2\ZZ+1) \}$ and $\cE_2=\{(x,y)\in\cE:x\in(2\ZZ+1),~y\in2\ZZ \}$. Since $\det A=1=da-bc=dx-by=mx+ny$, we have that $d=m$ and $b=-n$. We rule out when $v\in\cO,\cE_1$ since $c\neq 0 (\mod 2)$ in those cases. Consider $v\in\cE_2$. Then $x$ is odd and $y$ is even. $d$ must be odd since $mx+ny=1$ and $y$ is even. And $b$ must be even since $ad-bc=1$\\
\compav{finish this and say the proof for $\cE_1$ is analogous.}

\begin{lem}
Let $v_0\RR^2$. Then for any geodesic $\alpha$
\end{lem}

\subsection{empty}
\noindent Consider the following families of unit squares in $\RR^3$:\\
\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}

\subsection{Translation Surface Cover of $\bGs$}

$\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
